# CheckM2

## Introduction

CheckM2 is a basic secondary analysis to test the quality of the samples provided with a score for completeness and contamination [@chklovski2023].

## Methods

### CheckM2: slurm job

I used the script [run_checkm2TN.sh](https://github.com/tobiasnunn/tnunn_research/blob/main/Research_Experience_Module/00_scripts/Slurmscripts/run_checkm2TN.sh) to run CheckM2 on all 10 fasta files at once.

### CheckM2: R table generation

The CheckM2 job produced a [set of files](https://github.com/tobiasnunn/tnunn_research/tree/main/Research_Experience_Module/02_middle-analysis_outputs/CheckM2_20241228):

```{r checkm2_files}
#| echo: false
#| label: tbl-files
#| tbl-cap: "CheckM2 output files"

filepath <- "../02_middle-analysis_outputs/CheckM2_20241228"

# List files in a directory
files <- list.files(path = filepath, full.names = FALSE, recursive = TRUE)


file_info <- data.frame(
  Filename = files
)

# Display as a formatted table
knitr::kable(file_info)

```

The file I needed was `quality_report.tsv`, which gives info on each sample:

-   Completeness

-   Contamination

-   Completeness Model Used

-   Translation Table Used

I created an R script [CheckM2tablegenerator.R.](https://github.com/tobiasnunn/tnunn_research/blob/main/Research_Experience_Module/00_scripts/Rscripts/CheckM2tablegenerator.R) This produced @tbl-m2

Note: there were other outputs to CheckM2, I want to follow up on this during my dissertation, see @sec-genprog. Particularly the DIAMOND_RESULTS.tsv file looks interesting as it seems to have UNIPROT protein IDs linked to KEGG KO values, which is what I used eggNOG-mapper to do later on.

```{r quality_report}
#| echo: false
#| label: tbl-diamond
#| tbl-cap: "First five rows of DIAMOND_RESULTS.tsv file"


tsv_data <- read.delim("../02_middle-analysis_outputs/CheckM2_20241228/diamond_output/DIAMOND_RESULTS.tsv", sep = "\t", header = FALSE, nrows = 5)  # Preview first 5 rows

knitr::kable(tsv_data)

```

### Results

```{r checkm2, echo=FALSE, warning=FALSE}
#| echo: FALSE
#| warning: FALSE
#| label: tbl-m2
#| tbl-align: center
#| tbl-cap: "Quality analysis from CheckM2 based on samples taken from the skin microbiome of *Dendrobates tinctorius*"

library(tidyverse)
library(gt)

checkm2 <- read.csv("../03_final_outputs/metadata_and_quality_tables/checkm2.csv")
checkm2 %>%
  arrange(desc(Completeness)) %>%
  gt()  %>%
  tab_source_note(source_note = "Source: CheckM2 predict performed on 2024-12-28") %>%
  opt_row_striping() %>%
  opt_stylize(style = 5, color = "blue") %>%
  data_color(
    columns = Completeness,
    method = "numeric",
    palette = "Blues",
    domain = c(90, 100),
    reverse = FALSE
  ) %>%
   data_color(
    columns = Contamination,
    method = "numeric",
    palette = "Oranges",
    domain = c(0, 3),
    reverse = FALSE
  ) %>%
  cols_label(Name = "Accession",
             Completeness_Model_Used = "Completeness model used",
             Translation_Table_Used = "Translation table used",
             Additional_Notes = "Additional notes") %>%
  cols_width(
    Completeness_Model_Used ~ px(250)
  )
```

This is the table of CheckM2 analysis, there are many differences in which fields are produced, also, there appears to be differences in the values themselves for Completeness and Contamination, perhaps part of the comparitive analysis, done in the conclusion should be the date at which both were at, as older versions will probably have less accurate values, and this may cause the difference as genomics is a fast evolving field. A noteable difference is that CheckM used an internal diamond database, whereas CheckM2 required me to download my own.

### Scripts

-   [run_checkm2TN.sh](https://github.com/tobiasnunn/tnunn_research/blob/main/Research_Experience_Module/00_scripts/Slurmscripts/run_checkm2TN.sh)

-   [CheckM2tablegenerator.R.](https://github.com/tobiasnunn/tnunn_research/blob/main/Research_Experience_Module/00_scripts/Rscripts/CheckM2tablegenerator.R)

### Comparison of CheckM and CheckM2

```{r m_vs_m2, echo=FALSE, warning=FALSE}
#| echo: FALSE
#| label: tbl-mm2
#| tbl-align: center
#| tbl-cap: "Comparison of quality metrics for the same fasta files between CheckM/1.1.3 and CheckM2/0.1.3"

#----------------------combi table gen-------------------------------------
checkm <- read.csv("../03_final_outputs/metadata_and_quality_tables/checkm.csv")
checkm2 <- read.csv("../03_final_outputs/metadata_and_quality_tables/checkm2.csv")
names(checkm2)[names(checkm2) == "Name"] <- "accession"
combicheck <- merge(checkm, checkm2, by = "accession", all = TRUE)
#----------------------combi table product---------------------------------
columnnames <- c("accession", "Completeness.x", "Contamination.x", "Completeness.y", "Contamination.y")
combicheck[columnnames] %>%
  arrange(accession) %>%
  gt() %>%
  fmt_number(
    columns = c(Completeness.x, Contamination.x, Completeness.y, Contamination.y),
    decimals = 2,
    use_seps = TRUE)  %>%
  tab_source_note(source_note = md("Source:  
                  CheckM lineage_wf performed on 2024-12-25  
                  CheckM2 predict performed on 2024-12-28")) %>%
  opt_row_striping() %>%
  opt_stylize(style = 5, color = "blue") %>%
  data_color(
    columns = c(Completeness.x, Completeness.y),
    method = "numeric",
    palette = "Blues",
    domain = c(90, 100),
    reverse = FALSE
  ) %>%
   data_color(
    columns = c(Contamination.x, Contamination.y),
    method = "numeric",
    palette = "Oranges",
    domain = c(0, 3),
    reverse = FALSE
  ) %>%
 # tab_header(
  #  title = "Table 5 - Comparison of quality metrics",
   # subtitle = md("for the same fasta files between CheckM/1.1.3 and CheckM2/0.1.3")
    # ) %>%
  tab_spanner(
    label = "CheckM/1.1.3",
    columns = c("Completeness.x", "Contamination.x")) %>%
  tab_spanner(
    label = "CheckM2/0.1.3",
    columns = c("Completeness.y", "Contamination.y")) %>%
  cols_label(Completeness.x = "Completeness",
         Completeness.y = "Completeness",
         Contamination.x = "Contamination",
         Contamination.y = "Contamination",
         accession = "Accession")
   
```
