---
date: "`r format(Sys.time(), '%B %d, %Y')`"
title: "Oxygen uptake rates in a typical gill-breather"
format: docx
editor: visual
---

## Practical 1: Assessment

## BSX-2030 Integrated Zoology

## Student number: 500690571

![© Hans Hillewaert [CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)](1024px-Carcinus_maenas.jpg){fig-align="center"}

{{< pagebreak >}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(flextable)
library(equatags)
library(officer)
```

## Task 1

Task 1: Present the data in 4 tables i.e. Tables 6, 7, 8 and 9 from the practical handout. Each table needs a caption. Table titles go above the table

```{r rawdata, echo=FALSE}

crab_data <- data.frame(crab_name = c("A1", "A2", "A1", "A2", "A1", "A2"),
                        treatment = c("Submerged", "Submerged", "Aerial Exposure", "Aerial Exposure", "Resubmerged","Resubmerged"),
                        po_diff = c(13.3, 14.3, -0.9, 1, 15.2, 7.8),
                        time_diff = c(19, 19, 61, 61, 15, 15),
                        mass = c(0.02691, 0.02773,0.02691, 0.02773,0.02691, 0.02773),
                        volume = c(795,800,795,800,795,800)) 
# then add calculated columns
crab_data$solubility <-  crab_data$po_diff * 0.0000387
crab_data$solubility_by_volume <- crab_data$solubility * crab_data$volume
crab_data$solubility_by_time <- crab_data$solubility_by_volume * (60 / crab_data$time_diff)
crab_data$oxygen_by_mass <- crab_data$solubility_by_time / crab_data$mass
```

### Table 6 - calculation of Oxygen uptake for crabs in two aqueous conditions

```{r table6, echo=FALSE}

# bring in data
table6_data <- filter(crab_data, !treatment == "Aerial Exposure") %>%
  mutate(across(-c(crab_name, treatment, solubility), \(x) round(x, 2))) %>%
  select(-c(mass, oxygen_by_mass, volume))
table6 <- flextable(table6_data )

#headers and content
table6 <- set_header_labels(table6,
                            crab_name = "Crab\nName",
                            treatment = "Treatment",
                            time_diff = "Time\nbetween\nreadings\n(min)") %>%
  mk_par(
    # this is column A
    part = "header", j = "po_diff",
    value = as_paragraph("Difference\nin pO", as_sub("2"), "\n", "(mmHg)")
  ) %>% 
  mk_par(
    # this is column C
    part = "header", j = "solubility",
    value = as_paragraph("Difference\nin pO", as_sub("2"), "\n", "in water (mmHg)") 
  ) %>% 
  mk_par(
    # this is column D
    part = "header", j = "solubility_by_volume",
    value = as_paragraph("Difference\nin pO", as_sub("2"), "\naccounting\nfor volume\n(ml)") 
  ) %>% 
  mk_par(
    # this is column E
    part = "header", j = "solubility_by_time",
    value = as_paragraph("Oxygen Uptake\n(ml", as_sup("-1"), " O", as_sub("2"), " h", as_sup("-1"), ")"))
table6 <- theme_vanilla(table6)
table6 <- add_footer_lines(table6, "Data collected on 2025-01-30 at Deiniol Road, Brambell Building, 1st Floor Lab B1")
table6 <- color(table6, part = "footer", color = "#666666")
# table6 <- set_caption(table6, caption = paste0("New York Air Quality Measurements"))
table6 <- add_header_row(table6,
  colwidths = c(2, 1, 1, 1, 1, 1),
  values = c("", "A", "B", "C", "D","E")
)

#footnotes
table6 <- set_table_properties(table6, width = 1, layout = "autofit") %>%
footnote(i = 2, j = 5, 
           part = "header",
           ref_symbols = "*", 
         value = as_paragraph("value from column A ", as_equation("\\times"), " solubility coefficient (3.87", as_equation("\\times"), " 10", as_sup("-5"), ")")) %>%
footnote(i = 2, j = 6, 
           part = "header",
           ref_symbols = "†", 
         value = as_paragraph("value from column C ", as_equation("\\times"), " vol of water (ml)")) %>%
footnote(i = 2, j = 7, 
           part = "header",
           ref_symbols = "‡", 
         value = as_paragraph("value from column D " , as_equation("\\times"), " (", as_equation(x = "{60 \\over B}"), ") (ml", as_sup("-1"), " O", as_sub("2"), " h", as_sup("-1"), ")"))

table6 <- table6 %>% 
  set_formatter(solubility = function(x) {
    formatC(x, format = "e", digits = 2)
  })

table6 <- fontsize(table6, size = 10, part = "header")
# poss options for format 
# e = "e-5" / f = "0".00" / E = "E-5" / g = normal / G = normal but no trailing 0s # / fg apparantly does same as G / s is weird 
table6
```

### Table 7 - Oxygen uptake rate accounting for weight and converting to mmol for shore crabs in two aqueous conditions

```{r table7, echo=FALSE}
table7_data <- filter(crab_data, !treatment == "Aerial Exposure") %>% 
  mutate(column_c = oxygen_by_mass / 22.414) %>%
  select(crab_name, treatment, solubility_by_time, oxygen_by_mass, column_c) %>%
  mutate(across(-c(crab_name, treatment), \(x) round(x, 2)))
# headers
table7 <- flextable(table7_data)
table7 <- set_header_labels(table7,
                            crab_name = "Crab\nName",
                            treatment = "Treatment") %>% 
  mk_par(
    # column A
    part = "header", j = "solubility_by_time",
    value = as_paragraph("Oxygen uptake rate\n(ml", as_sup("-1"), " O", as_sub("2"), " h", as_sup("-1"), ")")
  ) %>%
  mk_par(
    # Column B
    part = "header", j = "oxygen_by_mass",
    value = as_paragraph("Oxygen uptake\nrate per unit mass\n(ml O", as_sub("2"), " kg",  as_sup("-1"), " h", as_sup("-1"), ")")
  ) %>%
  mk_par(
    # column C
    part = "header", j = "column_c",
    value = as_paragraph("Standardised\nOxygen uptake rate\n(mmol O", as_sub("2"), " kg",  as_sup("-1"), " h", as_sup("-1"), ")")
  )

# theme stuff
table7 <- theme_vanilla(table7)
table7 <- add_footer_lines(table7, "Data collected on 2025-01-30 at Deiniol Road, Brambell Building, 1st Floor Lab B1")
table7 <- color(table7, part = "footer", color = "#666666")
table7 <- set_caption(table7, caption = "God knows")
table7 <- add_header_row(table7,
  colwidths = c(2, 1, 1, 1),
  values = c("", "A", "B", "C")
)
# footnotes and extra
table7 <- set_table_properties(table7, width = 1, layout = "autofit") %>%
footnote(i = 2, j = 4, 
           part = "header",
           ref_symbols = "*", value = as_paragraph("value from column A divided by body mass in kg")) %>%
footnote(i = 2, j = 5, 
           part = "header",
           ref_symbols = "†", value = as_paragraph("value from column B divided by 22.414 (the conversion ratio between moles and litres). Conversion to mmol is in aid of comparing these values with those calculated in air (Table 8)")) 
table7 <- fontsize(table7, size = 10, part = "header")
table7

```

### Table 8 - Calculating Oxygen uptake for shore crabs exposed to air

```{r table8, echo=FALSE}
# can we avoid mutating the columns in thedata frame? this should bea display feature
# mutate(across(-c(crab_name, treatment), ~ signif(.x, digits = 2)))
# data stuff
aerial_data <- filter(crab_data, treatment == "Aerial Exposure") %>% 
  mutate(rt = 17969) %>%
  mutate(vol_of_oxy = (po_diff/rt) * (volume/1000)) %>%
  mutate(div_weight = vol_of_oxy / mass) %>%
  mutate(per_hour = div_weight * (60/61)) %>%
  select(crab_name, treatment, rt, po_diff, vol_of_oxy, div_weight, per_hour)

table8 <- flextable(aerial_data )
# headers
table8 <- set_header_labels(table8,
                            crab_name = "Crab\nName",
                            treatment = "Treatment") %>% 
  mk_par(
    # this is column A
    part = "header", j = "rt",
    value = as_paragraph("Universal gas\nconstant", as_equation("\\times"), "assumed\ntemperature\n(", as_equation("RT"), ")")
  ) %>%
  mk_par(
    # this is column B
    part = "header", j = "po_diff",
    value = as_paragraph("Difference\nin pO", as_sub("2"), "\n", "(mmHg)")
  ) %>%
  mk_par(
    # this is column C
    part = "header", j = "vol_of_oxy",
    value = as_paragraph("Difference\nin pO", as_sub("2"), "\n(mmol O", as_sub("2"),")")
  ) %>%
  mk_par(
    # this is D
    part = "header", j = "div_weight",
    value = as_paragraph("Difference\nper unit\nmass\n(mmol O", as_sub("2"), " kg",  as_sup("-1"), ")")
  ) %>%
  mk_par(
    # this is column E
    part = "header", j = "per_hour",
    value = as_paragraph("Oxygen uptake rate\n(mmol O", as_sub("2"), " kg",  as_sup("-1"), " h", as_sup("-1"), ")")
  )
table8 <- theme_vanilla(table8)
table8 <- add_footer_lines(table8, "Data collected on 2025-01-30 at Deiniol Road, Brambell Building, 1st Floor Lab B1")
table8 <- color(table8, part = "footer", color = "#666666")
table8 <- set_caption(table8, caption = "Crabs in the air")
table8 <- add_header_row(table8,
  colwidths = c(2, 1, 1, 1, 1, 1),
  values = c("", "A", "B", "C", "D", "E")
)

# footers and extra
table8 <- set_table_properties(table8, width = 1, layout = "autofit") %>%
footnote(i = 2, j = 3, 
           part = "header",
           ref_symbols = "*", 
         value = as_paragraph("The universal gas constant is 62.36, the experiment assumes a temperature of 15\u00b0C (which is 288.15\u00b0K)")) %>%
footnote(i = 1, j = 4, 
           part = "body",
           ref_symbols = "†", 
         value = as_paragraph("The container of crab A1 was found to have increased in partial pressure of Oxygen between readings, which would theoretically mean the crab lost oxygen, it is more likely this is an anomaly.")) %>%
footnote(i = 2, j = 5,
           part = "header",
           ref_symbols = "‡",
         value = as_paragraph(as_equation("{B \\over A} \\times"), " volume of air in litres ", "(mmol O", as_sub("2"),"). This reflects the equation for the ideal gas law, rearranged to find", as_equation("n"), " (to find the quantity of moles of oxygen), ", as_equation("{P \\over RT} \\times V = n"))) %>%
footnote(i = 2, j = 6,
           part = "header",
           ref_symbols = "§",
         value = as_paragraph("The oxygen difference in millimoles divided by the mass, in kg, for each crab (0.795 for A1 and 0.800 for A2). It should be noted that this step is not in the handout, but I felt it pertinent to include as the values of Oxygen uptake rate for the aqueous samples are in (mmol O", as_sub("2"), " kg",  as_sup("-1"), " h", as_sup("-1"), ")")) %>%
footnote(i = 2, j = 7,
           part = "header",
           ref_symbols = "¶",
         value = as_paragraph("The value in column C ", as_equation("\\times {60 \\over time interval}"), "(mmol O", as_sub("2"), " kg",  as_sup("-1"), " h", as_sup("-1"), "). Time interval recorded for both as 61 minutes"))


table8 <- table8 %>% style(
    j = c("po_diff", "vol_of_oxy", "div_weight", "per_hour"),
    i = ~ po_diff < 0 | vol_of_oxy < 0 | div_weight < 0 | per_hour < 0,
    pr_t = fp_text(color = "red"))

table8 <- set_formatter(
  x = table8,
  vol_of_oxy = function(x) {
    formatC(x, format = "e", digits = 2)
  },
  per_hour = function(x) {
    formatC(x, format = "e", digits = 2)
  },
  div_weight = function(x) {
    formatC(x, format = "e", digits = 2)
  }
)

table8 <- fontsize(table8, size = 9, part = "header")

table8

```

### Table 9 - Rates of oxygen uptake expressed as mmol O~2~  kg^-1^ h^-1^ in crabs submerged in seawater, exposed to roughly one hour aerial exposure and then resubmerged

``` {r table9, echo=FALSE}
table9 <- crab_data %>% select(
  "crab_name", "treatment"
) 

table9 <- table9 %>% left_join(table7_data, by = join_by(crab_name, treatment))

```
